A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-05-19 main.tex
Wrote on contexts.
Writing typing rules.
Writing details of typing rules.

\newcommand{\LTP}{$\lambda^{\triangleright\%}$}
\newcommand{\LMD}{$\lambda^{\textrm{MD}}$}
\newcommand{\LLF}{$\lambda^{\textrm{LF}}$}

@@

\newcommand{\TVar}{\textsc{T-Var}}
\newcommand{\TAbs}{\textsc{T-Abs}}
\newcommand{\TApp}{\textsc{T-App}}
\newcommand{\TConv}{\textsc{T-Conv}}
\newcommand{\TTB}{\textsc{T-$\TB$}}
\newcommand{\TTBL}{\textsc{T-$\TBL$}}
\newcommand{\TGen}{\textsc{T-Gen}}
\newcommand{\TIns}{\textsc{T-Ins}}
\newcommand{\TCsp}{\textsc{T-Csp}}

@@

\begin{align*}
    \textrm{Type variables} && & X,Y,Z \\
    \textrm{Constants} && & c \\
    \textrm{Variables} && & x,y,z \\
    \textrm{Stage variables} && & \alpha,\beta,\gamma \\
    \textrm{Stage} && & A,B,C \\
    \textrm{Signature} && \Sigma & ::= \phi \mid X::K \mid c:\tau \\
    \textrm{Kinds} && K,J,I,H,G & ::= * \mid \Pi x:\tau.K\\
    \textrm{Types} && \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
    \textrm{Terms} && M,N,L,O,P & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M 
    \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M\\ 
    \textrm{Contexts} && \Gamma & ::= \phi \mid  \Gamma,x:\tau@A\ (x\not\in\textrm{FV}(\G)) \\
\end{align*}

@@

@@

% 自由変数

As usual, the variable $x$ is bound in $\lambda x:\tau.M$
and the transition variable $\alpha$ is bound in $\Lambda \alpha.M$.
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.
We omit their definitions.

@@

\begin{definition}[Reduction]
There are three reduction rules ($\longrightarrow_\beta, \longrightarrow_\blacklozenge, \longrightarrow_\Lambda$) in \LMD.
Congruence rules which are omitted from the defnition.
\begin{align*}
    & (\lambda x:\tau.M) N \longrightarrow_\beta M[x \mapsto N] \\
    % & (\Pi x:\tau.\sigma) M \longrightarrow_\gamma \sigma[x \mapsto M] \\
    & \TBL_\alpha \TB_\alpha M \longrightarrow_\blacklozenge M \\
    & (\Lambda \alpha.M)\ \epsilon \longrightarrow_\Lambda M[\alpha \mapsto \epsilon]
\end{align*}
We write $ M \longrightarrow M'$ iff $ M \longrightarrow_\beta M'$, $ M \longrightarrow_\blacklozenge M'$, or $ M \longrightarrow_\Lambda M'$.
\end{definition}

@@

\subsection{Type System}
\subsubsection{Type Equivalence Rules}

>>>>

\newcommand{\LTP}{$\lambda^{\triangleright\%}$\ }
\newcommand{\LMD}{$\lambda^{\textrm{MD}}$\ }
\newcommand{\LLF}{$\lambda^{\textrm{LF}}$\ }

@@

\newcommand{\TConst}{\textsc{T-Const}}
\newcommand{\TVar}{\textsc{T-Var}}
\newcommand{\TAbs}{\textsc{T-Abs}}
\newcommand{\TApp}{\textsc{T-App}}
\newcommand{\TConv}{\textsc{T-Conv}}
\newcommand{\TTB}{\textsc{T-$\TB$}}
\newcommand{\TTBL}{\textsc{T-$\TBL$}}
\newcommand{\TGen}{\textsc{T-Gen}}
\newcommand{\TIns}{\textsc{T-Ins}}
\newcommand{\TCsp}{\textsc{T-Csp}}

@@

\begin{align*}
    \textrm{Type variables} && & X,Y,Z \\
    \textrm{Constants} && & c \\
    \textrm{Variables} && & x,y,z \\
    \textrm{Stage variables} && & \alpha,\beta,\gamma \\
    \textrm{Stage} && & A,B,C \\
    \textrm{Kinds} && K,J,I,H,G & ::= * \mid \Pi x:\tau.K\\
    \textrm{Types} && \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
    \textrm{Terms} && M,N,L,O,P & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M 
    \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M\\ 
    \textrm{Signature} && \Sigma & ::= \phi \mid X::K \mid c:\tau \\
    \textrm{Contexts} && \Gamma & ::= \phi \mid  \Gamma,x:\tau@A\ (x\not\in\textrm{FV}(\G)) \\
\end{align*}

@@

% Contexts
Contexts are sequences of triplexes of a variable, its type, and its stage.
In order to restrict contexts to only well-formed ones, there is a condition of $x\notin\FV(\G)$.
Because of this condition, there is no type with free variables in contexts.

@@

% 自由変数
As usual, the variable $x$ is bound in $\lambda x:\tau.M$
and the transition variable $\alpha$ is bound in $\Lambda \alpha.M$.
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.
We omit their definitions.

@@

\begin{definition}[Reduction]
There are three reduction rules ($\longrightarrow_\beta, \longrightarrow_\blacklozenge, \longrightarrow_\Lambda$) in \LMD.
Congruence rules which are omitted from the definition.
\begin{align*}
    & (\lambda x:\tau.M) N \longrightarrow_\beta M[x \mapsto N] \\
    % & (\Pi x:\tau.\sigma) M \longrightarrow_\gamma \sigma[x \mapsto M] \\
    & \TBL_\alpha \TB_\alpha M \longrightarrow_\blacklozenge M \\
    & (\Lambda \alpha.M)\ \epsilon \longrightarrow_\Lambda M[\alpha \mapsto \epsilon]
\end{align*}
We write $ M \longrightarrow M'$ iff $ M \longrightarrow_\beta M'$, $ M \longrightarrow_\blacklozenge M'$, or $ M \longrightarrow_\Lambda M'$.
\end{definition}

@@

\subsection{Type System}

% 総論

In this section, we define the type system of \LMD.
Especially, we discuss on the type equality of \LMD in detail
because it is the main contribution of this paper.
\LMD is developed on the basis of \LTP but there is no type equivalence rules in \LTP.

\begin{definition}[Typing]
    The typing relation $ \G \V M : \tau @ A $ is the least relation closed under the rules in Figure \ref{fig:typing-rules}.
\end{definition}

% 各論
The rules \TVar, \TAbs, and \TApp are almost same as those in simply typed lambda calculus if you ignore stage annotations.

\begin{figure}
\begin{center}
    \infrule{c:\tau \in \Sigma \andalso \G\V \tau::*@A}{\G \V c:\tau@A}{\TConst} \andalso
    \infrule{x:\tau@A \in \G \andalso \G\V \tau::*@A}{\G \V x:\tau@A}{\TVar} \\[2mm]
    \infrule{\G\V \sigma::*@A\andalso\G,x:\sigma@A\V M:\tau@A}{\G\V(\lambda (x:\sigma).M):(\Pi (x:\sigma).\tau)@A}{\textsc{T-Abs}} \\[2mm]
    \infrule{\G\V M:(\Pi (x:\sigma).\tau)@A \andalso \G\V N:\sigma@A}{\G\V M\ N : \tau[x\mapsto N]@A}{\textsc{T-App}} \andalso
    \infrule{\G\V M:\tau@A \andalso \G\V \tau\equiv \sigma :: K@A}{\G\V M:\sigma@A}{\textsc{T-Conv}} \\[2mm]
    \infrule{\G\V M:\tau@{A\alpha}}{\G\V\TB_{\alpha}M:\TW_{\alpha}\tau@A}{\textsc{T-$\TB$}} \andalso
    \infrule{\G\V M:\TW_{\alpha}\tau@A}{\G\V\TBL_{\alpha}M:\tau@{A\alpha}}{\TTBL} \\[2mm]
    \infrule{\G\V M:\tau@A \andalso \alpha\notin\rm{FTV}(\G)\cup\rm{FTV}(A)}{\G\V\Lambda\alpha.M:\forall\alpha.\tau@A}{\textsc{T-Gen}} \andalso
    \infrule{\G\V M:\forall\alpha.\tau@A}{\G\V M\ \epsilon:\tau[\alpha \mapsto \epsilon]@A}{\textsc{T-Ins}} \andalso
    \infrule{\G\V M:\tau@A}{\G\V \%_\alpha M:\tau@{A\alpha}}{\textsc{T-Csp}} \andalso
    \caption{Typing Rules}
    \label{fig:typing-rules}
\end{center}
\end{figure}

\subsubsection{Type Equivalence Rules}
